\bf Mechanistic models for panel data: analysis of ecological experiments with four interacting species

### 2024-04-03 ms.Rnw
Update

\begin{figure}[H]
\centering
<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=16, height=10>>=  
   mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
     cbind(mixed_nopara_data[[i]], group=paste("Trial", trails[i]))
   }))
    mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
     cbind(mixed_para_data[[i]], group=paste("Trial", trails[i]))
   }))

p1<- ggplot(mixed_nopara_data_combined, aes(x=day)) + 
   geom_line(aes(y=dent.adult, group=interaction(group, "dent.adult"))) + 
   geom_line(aes(y=lum.adult, color="blue", group=interaction(group, "lum.adult"))) +
   scale_x_continuous(limits = c(0, 52)) +
   scale_y_continuous(limits = c(0, 175)) +
   facet_grid(. ~ group, scales = "fixed", space = "free") +
   theme(
     strip.text = element_text(size=10),
     axis.text.x=element_text(color="black"),
     axis.ticks.x=element_line(color="black"),
     axis.text.y=element_text(color="black"),
     axis.ticks.y=element_line(color="black"),
     axis.title.x=element_blank(),
     axis.title.y=element_blank(),
     legend.position="none",
     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
     strip.background = element_blank(), # Optional: Remove facet label background
     strip.placement = "outside"
   ) +
   labs(title="Daphnia density in Non-parasitized Dynamics") +
   guides(colour=guide_legend(title="Species"))

p2<- ggplot(mixed_para_data_combined, aes(x=day)) + 
   geom_line(aes(y=dentadult, group=interaction(group, "dentadult"))) + 
   geom_line(aes(y=lumadult, color="blue", group=interaction(group, "lumadult"))) +
    geom_line(aes(y=luminf,color="green", group=interaction(group, "luminf"))) + 
   geom_line(aes(y=dentinf, color="red", group=interaction(group, "dentinf"))) +

   scale_x_continuous(limits = c(0, 52)) +
   scale_y_continuous(limits = c(0, 175)) +
   facet_grid(. ~ group, scales = "fixed", space = "free") +
   theme(
     strip.text = element_text(size=10),
     axis.text.x=element_text(color="black"),
     axis.ticks.x=element_line(color="black"),
     axis.text.y=element_text(color="black"),
     axis.ticks.y=element_line(color="black"),
     axis.title.x=element_blank(),
     axis.title.y=element_blank(),
     legend.position="none",
     panel.spacing = unit(0, "lines"), # Adjust spacing between panels
     strip.background = element_blank(), # Optional: Remove facet label background
     strip.placement = "outside"
   ) +
   labs(title="Daphnia density in Parasitized Dynamics") +
   guides(colour=guide_legend(title="Species"))
grid.arrange(p1, p2, ncol = 1)
@
\caption{Plot of density of hosts in dynamics including both native and invasive host species of \textit{Daphnia}. The trail A to J represents different buckets in different experiment. The black lines represent the \textit{D. dentifera} density, the red lines show the \textit{D. lumholtzi} density, the green lines are the infected \textit{D. dentifera} density, and the blue lines indicate the infected \textit{D. lumholtzi} density.}
\end{figure}

>>>>

\begin{figure}[H]
\centering
<<echo=FALSE, fig.cap="Plot1", fig.align='center', fig=TRUE, width=16, height=10>>=
 mixed_nopara_data_combined <- do.call(rbind, lapply(1:length(mixed_nopara_data), function(i) {
  cbind(mixed_nopara_data[[i]], Trial = paste("Trial", trails[i]))
}))

# Combine parasitized data with trial information
mixed_para_data_combined <- do.call(rbind, lapply(1:length(mixed_para_data), function(i) {
  cbind(mixed_para_data[[i]], Trial = paste("Trial", trails[i]))
}))

# Assuming 'mixed_nopara_data_combined' is already your data frame as shown
# Create a mapping from Trial letters to bucket numbers
trial_to_bucket <- function(trial) {
  # Extract the letter from the trial name
  letter <- substr(trial, 7, 7)
  # Convert the letter to a bucket number (A=1, B=2, ..., J=10)
  bucket_number <- match(letter, LETTERS[1:26])
  # Return the bucket number prefixed with "Bucket"
  return(paste("Bucket", bucket_number))
}

# Apply the function to create a new 'Bucket' column
mixed_nopara_data_combined$Bucket <- sapply(mixed_nopara_data_combined$Trial, trial_to_bucket)
mixed_para_data_combined$Bucket <- sapply(mixed_para_data_combined$Trial, trial_to_bucket)
p1 <- ggplot(mixed_nopara_data_combined, aes(x = day)) +
  geom_line(aes(y = dent.adult, group = interaction(Bucket, "dent.adult"), color = "Dent Adult"), size = 1) +  # Map color to a factor for legend
  geom_line(aes(y = lum.adult, group = interaction(Bucket, "lum.adult"), color = "Lum Adult"), size = 1) +  # Map color to a factor for legend
  scale_x_continuous(limits = c(0, 52)) +
  scale_y_continuous(limits = c(0, 175)) +
  facet_grid(. ~ Bucket, scales = "fixed", space = "free") +
  theme_minimal() +
  labs(title = "Daphnia Density in Non-parasitized Dynamics",
       color = "Species") +
  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue"),
                     labels = c("Dent Adult" = "Dent adult", "Lum Adult" = "Lum adult")) +
  theme(legend.position = "right")  # Adjust the legend position as needed

p2 <- ggplot(mixed_para_data_combined, aes(x = day)) +
  geom_line(aes(y = dentadult, color = "Dent Adult", group = interaction(Bucket, "dentadult"))) +  # Added color mapping for dentadult
  geom_line(aes(y = lumadult, color = "Lum Adult", group = interaction(Bucket, "lumadult"))) +  # Ensure consistent color mapping
  geom_line(aes(y = luminf, color = "Lum Inf", group = interaction(Bucket, "luminf"))) +  # Ensure consistent color mapping
  geom_line(aes(y = dentinf, color = "Dent Inf", group = interaction(Bucket, "dentinf"))) +  # Ensure consistent color mapping
  scale_x_continuous(limits = c(0, 52)) +
  scale_y_continuous(limits = c(0, 175)) +
  facet_grid(. ~ Bucket, scales = "fixed", space = "free") +
  theme_minimal() +
  labs(title = "Daphnia Density in Parasitized Dynamics", color = "Species") +
  scale_color_manual(values = c("Dent Adult" = "black", "Lum Adult" = "blue", "Lum Inf" = "green", "Dent Inf" = "red"),
                     labels = c("Dent Adult" = "Dent adult", "Lum Adult" = "Lum adult", "Lum Inf" = "Lum inf", "Dent Inf" = "Dent inf")) +
  guides(color = guide_legend(title = "Species"))  # Ensure the legend is properly configured

# Arrange plots vertically
grid.arrange(p1, p2, ncol = 1)

@
\caption{Plot of density of hosts in dynamics including both native and invasive host species of \textit{Daphnia}. The trail A to J represents different buckets in different experiment. The black lines represent the \textit{D. dentifera} density, the red lines show the \textit{D. lumholtzi} density, the green lines are the infected \textit{D. dentifera} density, and the blue lines indicate the infected \textit{D. lumholtzi} density.}
\end{figure}
